Coherent optical creation of a single molecule

### 2020-12-03 paper.tex
Regenerate PDF, whitespace
Update on Overleaf.

Diverse species of fully quantum controlled  ultracold molecules are desired
for a wide variety of applications including precision measurements~\cite{
  Kondov2019,Nick_and_Ivan2017, PhysRevA.101.042504, Andreev2018,
  PhysRevLett.119.153001, hudson2011},
quantum simulations~\cite{Micheli2006, Yao2018, Wall2015, wall2015realizing},
quantum information processing~\cite{DeMille2002, Ni2018, Hudson2018, Lin2019},
and studies of ultracold chemistry~\cite{Bala2016,Hu1111,Segev2019,deJongh626}.
While many innovative approaches demonstrated in the last few years
have directly cooled different species of diatomic or polyatomic molecules
below 1~mK~\cite{Norrgard2016,Anderegg2018, Mitra1366, PhysRevX.10.021049,
  PhysRevLett.121.013202, Truppe2017},
the highest phase-space-density gas~\cite{Demarco2018} and
the coldest individual molecules~\cite{Zhang2020,He331}
have been achieved through the association of ultracold atoms.
% (any place to cite molecular ion work?)
% Citations are not complete. Add them as you see fit.

@@

The essence of an optical Raman transfer can be illustrated
using a three-level system shown in Fig.~\ref{f-theory}a,
where the initial atomic state and the target weakly bound molecular state are coupled to an intermediate state by two lasers with Rabi frequencies, $\Omega_a$ and $\Omega_m$, with one-photon detuning $ \Delta $, and two-photon detuning, $ \delta$.  %(Ultimately, we calculate with contribution of all vibrational states of the excited electronic potential, $\mathrm{c^3\Sigma^+}$)
The transfer Raman Rabi Rate is given by $\Omega_a\Omega_m / 2\Delta$~\cite{Wineland2003}.
% , is accompanied by a photon scattering rate $\Gamma_e (\Omega_a^2 + \Omega_m^2)/4\Delta^2$
% $\Gamma_e \Omega^2 / 4\Delta^2 $, where $ \Gamma_e $ is the excited state linewidth
Unlike Raman transitions in atoms, the two Rabi frequencies are greatly imbalanced %. Since $ \Omega_m $ is generally much larger than $ \Omega_a $
due to the small wavefunction overlap between the atomic state and the intermediate state, %the atomic state and the excited molecular state,
and therefore the scattering is predominantly from the target molecular state. Furthermore, the energy difference between the atomic state and target molecular state is small ($ < 1~\mathrm{GHz} $) compared to the single-photon detuning, $ \Delta $, so the target molecular state can scatter off both beams roughly equally. Thus, the scattering rate is given by $ \Gamma_e \Omega_m^2 / 2\Delta^2$, where $ \Gamma_e $ is the excited-state linewidth \footnote{We choose the two beams to have equal power, which gives the highest Raman Rabi rate at a fixed total power. Thus, this results in a simple factor of 2 coming from scattering off 2 beams.}.
The ratio between the Raman Rabi frequency and the scattering rate is therefore $ \Omega_a/\Omega_m \times \Delta/\Gamma_e $. %, depends on the ratio of the two matrix elements and how far detuned the laser is from the transition in units of the linewidth.
To ensure a coherent process, a detuning as large as possible, while maintaining a realistic Raman Rabi frequency, is preferred. %However, the detuning cannot be too large, since that will reduce the Raman Rabi frequency.

@@

In addition to the intermediate state,
choosing an initial and a final state for a large ratio $\Omega_a/\Omega_m$
maximizes $\Omega_R$ at a given detuning.
Furthermore, a larger ratio also relaxes the intensity stability requirement for the Raman lasers,
because this is also the ratio between $\Omega_R$
and the AC Stark shift of the molecular state, $\Omega_m^2 / 2\Delta$
\footnote{There is an additional factor of 2, with both beams at equal power,
  to account for the Stark shift caused by both beams.}.
Due to the small extent of the intermediate-state wavefunction
compared to that of the trapped atoms,
$\Omega_a$ is approximately proportional to
the amplitude of the relative atomic wavefunction at short distance,
within the range of the molecular potential.
To increase this amplitude, one can increase the external confinement of atom pairs.
Using a harmonic approximation,
the short-range amplitude is proportional to $ \omega_{\text{trap}}^{3/4} $ or $P^{3/8}$,
where $ \omega_{\text{trap}} $ is the trap frequency and $P$ is the optical power in the tweezer trap \cite{Mies2000}. However, additional power may not always be available, and will also lead to additional undesired scattering.
Alternatively, one can choose an atomic pair state with a large scattering length
(positive or negative).
For such states, the amplitude of the relative atomic wavefunction is substantially enhanced
at short range, as shown in Fig.~\ref{f-theory}c.
% The increase in the coupling is proportional to (quote/cite Olive's equation?).
For our system of Na and Cs atoms,
we choose a spin-state combination $\ket{\uparrow_{\Na} \downarrow_{\Cs}}\equiv \ket{f=2,m_f=2}_{\Na}\ket{f=3,m_f=3}_{\Cs}$ that has a large and negative scattering length of
$a(\uparrow_{\Na} \downarrow_{\Cs}) \approx -700a_0$~\cite{Hood2019}.
All other stable spin combinations give smaller scattering lengths ($<~50~a_0$).
% We denote the possible hyperfine states of the atoms as $ \ket{\uparrow_{\Cs}} = \ket{f=4,m_f=4}_{\Cs}, \ket{\downarrow_{\Cs}} = \ket{f=3,m_f=3}_{\Cs}, \ket{\uparrow_{\Na}} = \ket{f=2,m_f=2}_{\Na}, $ and $ \ket{\downarrow_{\Na}} = \ket{f=1,m_f=1}_{\Na}$. Among the stable spin combinations, $\ket{\uparrow_{\Na}\uparrow_{\Cs}}$ and $\ket{\downarrow_{\Na}\downarrow_{\Cs}} $ both have small scattering lengths of $ 30.4a_0 $, and $ 13.7a_0 $ respectively, but the $\ket{\uparrow_{\Na} \downarrow_{\Cs}} $ combination has a large and negative scattering length of $ a(\uparrow_{\Na} \downarrow_{\Cs}) = -693.8a_0 $ (interaction shift $\approx$ binding?)~\cite{Hood2019}.

@@

We perform the experiment at various tweezer powers to extract the power dependence of $ \Omega_m $ and $ \Omega_a $, as shown in Fig.~\ref{f-det}c. $ \Omega_m $ scales with $ P^{0.5} $ as expected, and we obtain a fit value of $2\pi\times180.87(7)~\mathrm{MHz}/\sqrt{\mathrm{mW}}$.
This number is within $40~\mathrm{\%}$
of the value $2\pi\times296 ~\mathrm{MHz}/\sqrt{\mathrm{mW}}$ calculated from theory. The scaling of $ \Omega_a $ differs from $ P^{0.5} $, which is the expected scaling from the laser intensity, due to the change in the atomic wavefunction caused by tighter confinement at higher power. For weakly interacting particles,
$\Omega_a$ scales as $P^{3/8}$.
However, due to the strong interaction between the two atoms in the $\ket{\uparrow_{\text{Na}}\downarrow_{\text{Cs}}}$ state, this approximation breaks down.
Instead, coupled-channel calculations show that the scaling
is very well approximated by $P^{0.29}$ within the range of confinement in our experiment. Thus, $ \Omega_a $ scales as $ P^{0.79} $, which fits well to our experimental result.
The fit value is $2\pi\times1.85(3)~\mathrm{MHz}/\mathrm{P^{0.79}}$. At $3.75~\mathrm{mW}$ tweezer power, we measured $ \Omega_m = 2\pi \times 348.3(3)~\mathrm{MHz}$ and $ \Omega_a = 2\pi\times 5.5(2)~\mathrm{MHz}$, resulting in a $ \Omega_m/\Omega_a$ ratio of about $ 63 $, which is close to the theory prediction of $77.7$~(Table~\ref{tab:rabi_freqs}).
Therefore, the matrix element ratio is not the cause of our lower than expected transfer efficiency.
Furthermore, we independently measure the natural linewidth of the $v'=0$ excited state to be no larger than $50~\mathrm{MHz}$ using photoassociation (PA) spectroscopy, which cannot explain the excess scattering either.
%We observe that the resonance frequency depends linearly on the tweezer power due to the differential light shift between the atomic and molecular state.
% From these measurements, we can extract a $\Omega_m$ of

%In order to calculate the ratio $\Omega_m/\Omega_a$,
%we now need to extract $ \Omega_a $.
%We do this by measuring the dependence of the Raman Rabi frequency on power and detuning.
% which depends on both $\Omega_m$ and $\Omega_a$.
%The Raman Rabi frequency shows a non-linear dependency on the tweezer power
%due to the change in the atomic wavefunction caused by
%tighter confinement at higher power, as shown in Fig.~\ref{f-det}b.

%Combined with the standard intensity factor, the Raman Rabi frequency should scale as $P^{1.29}$,
%which fits well to our experimental result, as shown in Fig.~\ref{f-det}b.
%Similar to the light shift, the detuning dependency of the Raman Rabi frequency
%is determined by a constant background component and
%a $v'=0$ component that scales as $1/\Delta$.
%The $v'=0$ component of the Raman Rabi frequency at our detuning is
%$2\pi\times1.10(2)~\mathrm{kHz\cdot mW^{-1.29}}$,
%or $2\pi \times 6.08(9)~\mathrm{kHz}$ at $3.75~\mathrm{mW}$ %tweezer power.
%Together with the $\Omega_m$ measured above, the Rabi frequency, $\Omega_a$, is
%$2\pi\times 5.5(2)~\mathrm{MHz}$.
%This gives a matrix-element ratio $\Omega_m/\Omega_a$ of $63(3)$,
%which is close to the theory prediction of $77.7$~(Table~\ref{tab:rabi_freqs}).

\begin{table}[ht]
  \centering
  \begin{tabular}{c|c|c}
    & Experiment & Theory \\ \hline
    $\Omega_R $ & $2\pi \times 3.28(4)~\mathrm{kHz}$ & $2\pi \times 15.3~\mathrm{kHz}$ \\
    $\Gamma_s$ & $2\pi \times 0.80(4)~\mathrm{kHz}$ & $ 2\pi \times 533~\mathrm{Hz}$ \\
    $\Omega_m$ & $2\pi \times 348.3(3)~\mathrm{MHz} $ & $2\pi \times 573 ~\mathrm{MHz}$ \\
    $\Omega_a$ & $2\pi \times 5.5(2)~\mathrm{MHz}$ & $2\pi \times 7.37 ~\mathrm{MHz}$ \\
    $\Omega_m/\Omega_a$ & $63(3)$ & $77.7$
  \end{tabular}
  \caption{Comparison between theory and experiment of $\Omega_R,\Gamma_s, \Omega_m, \Omega_a$. $\Omega_R$ and $ \Gamma_s $ are reported at $151~\mathrm{GHz}$ detuned from the $v' = 0$ state. All numbers are reported at $3.75$ mW tweezer power. }
  \label{tab:rabi_freqs}
\end{table}

%With the contribution of the $ v' = 0 $ state to the Raman Rabi frequency and scattering investigated,
We now consider the background effects from other states with larger single-photon detuning.
In the Raman Rabi frequency fit,
the fitted background is of opposite sign from the Raman Rabi frequency
for single-photon detunings red of the $v' = 0 $ transition.
Thus, this background reduces the Raman Rabi frequency by about $30~\mathrm{\%}$
at the current detuning.
However, this difference is not enough to explain the discrapancy of more than a factor of $10$
present in the experiment.
Due to the change in sign of the Raman Rabi frequency as a function of detuning
when crossing a resonance, the same background will increase the Raman Rabi frequency
for positive detunings from the $v' = 0$ transition.
Unfortunately, we observe additional nearby excited states
belonging to a different electronically excited state at higher frequencies
which prevent the blue side of the transition from being be usable for the Raman transition.

>>>>

Diverse species of fully quantum controlled  ultracold molecules are desired
for a wide variety of applications including precision measurements~\cite{
  Kondov2019,Nick_and_Ivan2017, PhysRevA.101.042504, Andreev2018,
  PhysRevLett.119.153001, hudson2011},
quantum simulations~\cite{Micheli2006, Yao2018, Wall2015, wall2015realizing},
quantum information processing~\cite{DeMille2002, Ni2018, Hudson2018, Lin2019},
and studies of ultracold chemistry~\cite{Bala2016,Hu1111,Segev2019,deJongh626}.
While many innovative approaches demonstrated in the last few years
have directly cooled different species of diatomic or polyatomic molecules
below 1~mK~\cite{Norrgard2016,Anderegg2018, Mitra1366, PhysRevX.10.021049,
  PhysRevLett.121.013202, Truppe2017},
the highest phase-space-density gas~\cite{Demarco2018} and 
%the coldest 
trapped individual molecules~\cite{Zhang2020,He331}
have been achieved through the association of ultracold atoms.
% (any place to cite molecular ion work?)
% Citations are not complete. Add them as you see fit.

@@

The essence of an optical Raman transfer can be illustrated
using a three-level system shown in Fig.~\ref{f-theory}a,
where the initial atomic state and the target weakly bound molecular state are coupled to an intermediate state by two lasers with Rabi frequencies, $\Omega_a$ and $\Omega_m$, with one-photon detuning $ \Delta $, and two-photon detuning, $ \delta$. Note that we use the convention where the Rabi frequency describes the population oscillation between the two states.  %(Ultimately, we calculate with contribution of all vibrational states of the excited electronic potential, $\mathrm{c^3\Sigma^+}$)
The transfer Raman Rabi Rate is given by $\Omega_a\Omega_m / (2\Delta)$~\cite{Wineland2003}.
% , is accompanied by a photon scattering rate $\Gamma_e (\Omega_a^2 + \Omega_m^2)/4\Delta^2$
% $\Gamma_e \Omega^2 / 4\Delta^2 $, where $ \Gamma_e $ is the excited state linewidth
Unlike Raman transitions in atoms, the two Rabi frequencies are greatly imbalanced %. Since $ \Omega_m $ is generally much larger than $ \Omega_a $
due to the small wavefunction overlap between the atomic state and the intermediate state, %the atomic state and the excited molecular state,
and therefore the scattering is predominantly from the target molecular state. Furthermore, the energy difference between the atomic state and target molecular state is small ($ < 1~\mathrm{GHz} $) compared to the single-photon detuning, $ \Delta $, so the target molecular state can scatter off both beams roughly equally. Thus, the scattering rate is given by $ \Gamma_e \Omega_m^2 / (2\Delta^2)$, where $ \Gamma_e $ is the excited-state linewidth \footnote{We choose the two beams to have equal power, which gives the highest Raman Rabi rate at a fixed total power. Thus, this results in a simple factor of 2 coming from scattering off 2 beams.}.
The ratio between the Raman Rabi frequency and the scattering rate is therefore $ \Omega_a/\Omega_m \times \Delta/\Gamma_e $. %, depends on the ratio of the two matrix elements and how far detuned the laser is from the transition in units of the linewidth.
To ensure a coherent process, a detuning as large as possible, while maintaining a realistic Raman Rabi frequency, is preferred. %However, the detuning cannot be too large, since that will reduce the Raman Rabi frequency.

@@

In addition to the intermediate state,
choosing an initial and a final state for a large ratio $\Omega_a/\Omega_m$
maximizes $\Omega_R$ at a given detuning.
Furthermore, a larger ratio also relaxes the intensity stability requirement for the Raman lasers,
because this is also the ratio between $\Omega_R$
and the AC Stark shift of the molecular state, $\Omega_m^2 / (2\Delta)$
\footnote{There is an additional factor of 2, with both beams at equal power,
  to account for the Stark shift caused by both beams.}.
Due to the small extent of the intermediate-state wavefunction
compared to that of the trapped atoms,
$\Omega_a$ is approximately proportional to
the amplitude of the relative atomic wavefunction at short distance,
within the range of the molecular potential.
To increase this amplitude, one can increase the external confinement of atom pairs.
Using a harmonic approximation,
the short-range amplitude is proportional to $ \omega_{\text{trap}}^{3/4} $ or $P^{3/8}$,
where $ \omega_{\text{trap}} $ is the trap frequency and $P$ is the optical power in the tweezer trap \cite{Mies2000}. However, additional power may not always be available, and will also lead to additional undesired scattering.
Alternatively, one can choose an atomic pair state with a large scattering length
(positive or negative).
For such states, the amplitude of the relative atomic wavefunction is substantially enhanced
at short range, as shown in Fig.~\ref{f-theory}c.
% The increase in the coupling is proportional to (quote/cite Olive's equation?).
For our system of Na and Cs atoms,
we choose a spin-state combination $\ket{\uparrow_{\Na} \downarrow_{\Cs}}\equiv \ket{f=2,m_f=2}_{\Na}\ket{f=3,m_f=3}_{\Cs}$ that has a large and negative scattering length of
$a(\uparrow_{\Na} \downarrow_{\Cs}) \approx -700a_0$~\cite{Hood2019}.
All other stable spin combinations give smaller scattering lengths ($<~50~a_0$).
% We denote the possible hyperfine states of the atoms as $ \ket{\uparrow_{\Cs}} = \ket{f=4,m_f=4}_{\Cs}, \ket{\downarrow_{\Cs}} = \ket{f=3,m_f=3}_{\Cs}, \ket{\uparrow_{\Na}} = \ket{f=2,m_f=2}_{\Na}, $ and $ \ket{\downarrow_{\Na}} = \ket{f=1,m_f=1}_{\Na}$. Among the stable spin combinations, $\ket{\uparrow_{\Na}\uparrow_{\Cs}}$ and $\ket{\downarrow_{\Na}\downarrow_{\Cs}} $ both have small scattering lengths of $ 30.4a_0 $, and $ 13.7a_0 $ respectively, but the $\ket{\uparrow_{\Na} \downarrow_{\Cs}} $ combination has a large and negative scattering length of $ a(\uparrow_{\Na} \downarrow_{\Cs}) = -693.8a_0 $ (interaction shift $\approx$ binding?)~\cite{Hood2019}.

@@

We perform the experiment at various tweezer powers to extract the power dependence of $ \Omega_m $ and $ \Omega_a $, as shown in Fig.~\ref{f-det}c. $ \Omega_m $ scales with $ P^{0.5} $ as expected, and we obtain a fit value of $2\pi\times180.87(7)~\mathrm{MHz}/\sqrt{\mathrm{mW}}$. The scaling of $ \Omega_a $ differs from $ P^{0.5} $, which is the expected scaling from the laser intensity, due to the change in the atomic wavefunction caused by tighter confinement at higher power. For weakly interacting particles,
$\Omega_a$ scales as $P^{3/8}$.
However, due to the strong interaction between the two atoms in the $\ket{\uparrow_{\text{Na}}\downarrow_{\text{Cs}}}$ state, this approximation breaks down.
Instead, coupled-channel calculations show that the scaling
is very well approximated by $P^{0.29}$ within the range of confinement in our experiment. Thus, $ \Omega_a $ scales as $ P^{0.79} $, which fits well to our experimental result.
The fit value is $2\pi\times1.85(3)~\mathrm{MHz}/\mathrm{P^{0.79}}$. At $3.75~\mathrm{mW}$ tweezer power, we measured $ \Omega_m = 2\pi \times 348.3(3)~\mathrm{MHz}$ and $ \Omega_a = 2\pi\times 5.5(2)~\mathrm{MHz}$, resulting in a $ \Omega_m/\Omega_a$ ratio of about $ 63 $, which is close to the theory prediction of $77.7$~(Table~\ref{tab:rabi_freqs}).
Therefore, the matrix element ratio is not the cause of our lower than expected transfer efficiency.
Furthermore, we independently measure the natural linewidth of the $v'=0$ excited state to be no larger than $50~\mathrm{MHz}$ using photoassociation (PA) spectroscopy, which cannot explain the excess scattering either.
% We observe that the resonance frequency depends linearly on the tweezer power due to the differential light shift between the atomic and molecular state.
% From these measurements, we can extract a $\Omega_m$ of

% In order to calculate the ratio $\Omega_m/\Omega_a$,
% we now need to extract $ \Omega_a $.
% We do this by measuring the dependence of the Raman Rabi frequency on power and detuning.
% which depends on both $\Omega_m$ and $\Omega_a$.
% The Raman Rabi frequency shows a non-linear dependency on the tweezer power
% due to the change in the atomic wavefunction caused by
% tighter confinement at higher power, as shown in Fig.~\ref{f-det}b.

% Combined with the standard intensity factor, the Raman Rabi frequency should scale as $P^{1.29}$,
% which fits well to our experimental result, as shown in Fig.~\ref{f-det}b.
% Similar to the light shift, the detuning dependency of the Raman Rabi frequency
% is determined by a constant background component and
% a $v'=0$ component that scales as $1/\Delta$.
% The $v'=0$ component of the Raman Rabi frequency at our detuning is
% $2\pi\times1.10(2)~\mathrm{kHz\cdot mW^{-1.29}}$,
% or $2\pi \times 6.08(9)~\mathrm{kHz}$ at $3.75~\mathrm{mW}$ %tweezer power.
% Together with the $\Omega_m$ measured above, the Rabi frequency, $\Omega_a$, is
% $2\pi\times 5.5(2)~\mathrm{MHz}$.
% This gives a matrix-element ratio $\Omega_m/\Omega_a$ of $63(3)$,
% which is close to the theory prediction of $77.7$~(Table~\ref{tab:rabi_freqs}).

% Theory numbers
% $2\pi \times 15.3~\mathrm{kHz}$
%  $ 2\pi \times 533~\mathrm{Hz}$
% $2\pi \times 573 ~\mathrm{MHz}$
% $2\pi \times 7.37 ~\mathrm{MHz}$

\begin{table}[ht]
  \centering
  \begin{tabular}{c|c}
    & Measured Values \\ \hline
    $\Omega_R $ & $2\pi \times 3.28(4)~\mathrm{kHz}$ \\
    $\Gamma_s$ & $2\pi \times 0.80(4)~\mathrm{kHz}$ \\
    $\Omega_m$ & $2\pi \times 348.3(3)~\mathrm{MHz} $  \\
    $\Omega_a$ & $2\pi \times 5.5(2)~\mathrm{MHz}$  \\
    $\Omega_m/\Omega_a$ & $63(3)$ 
  \end{tabular}
  \caption{Experimental values of $\Omega_R,\Gamma_s, \Omega_m, \Omega_a$. $\Omega_R$ and $ \Gamma_s $ are reported at $151~\mathrm{GHz}$ detuned from the $v' = 0$ state. All numbers are reported at $3.75$ mW tweezer power. }
  \label{tab:rabi_freqs}
\end{table}

% With the contribution of the $ v' = 0 $ state to the Raman Rabi frequency and scattering investigated,
We now consider the background effects from other states with larger single-photon detuning.
In the Raman Rabi frequency fit,
the fitted background is of opposite sign from the Raman Rabi frequency
for single-photon detunings red of the $v' = 0 $ transition.
Thus, this background reduces the Raman Rabi frequency by about $30~\mathrm{\%}$
at the current detuning.
However, this difference is not enough to explain the discrapancy of more than a factor of $10$
present in the experiment.
Due to the change in sign of the Raman Rabi frequency as a function of detuning
when crossing a resonance, the same background will increase the Raman Rabi frequency
for positive detunings from the $v' = 0$ transition.
Unfortunately, we observe additional nearby excited states
belonging to a different electronically excited state at higher frequencies
which prevent the blue side of the transition from being be usable for the Raman transition.
